Modal operators for a constructive account of well quasi-orders

### 2025-06-08 slides-types2025.tex
More work on the slides

>>>>

\documentclass[11pt,utf8,notheorems,compress,t,aspectratio=169]{beamer}
\usepackage{etex}

\usepackage{pgfpages}
\usepackage[export]{adjustbox}

% Workaround for the issue described at
% https://tex.stackexchange.com/questions/164406/beamer-using-href-in-notes.
\newcommand{\fixedhref}[2]{\makebox[0pt][l]{\hspace*{\paperwidth}\href{#1}{#2}}\href{#1}{#2}}

\usepackage[english]{babel}

\usepackage{mathtools}
\usepackage{xspace}
\usepackage{booktabs}
\usepackage{stmaryrd}
\usepackage{amssymb}
\usepackage{manfnt}
\usepackage{array}
\usepackage{ragged2e}
\usepackage{multicol}
\usepackage{tabto}
\usepackage{xstring}
\usepackage{proof}
\usepackage{agda}
\usepackage[all]{xy}
\xyoption{rotate}
\usepackage{tikz}
\usetikzlibrary{calc,shapes,shapes.callouts,shapes.arrows,patterns,fit,backgrounds,decorations.pathmorphing,positioning,svg.path}
\hypersetup{colorlinks=true}

\DeclareFontFamily{U}{bbm}{}
\DeclareFontShape{U}{bbm}{m}{n}
   {  <5> <6> <7> <8> <9> <10> <12> gen * bbm
      <10.95> bbm10%
      <14.4>  bbm12%
      <17.28><20.74><24.88> bbm17}{}
\DeclareFontShape{U}{bbm}{m}{sl}
   {  <5> <6> <7> bbmsl8%
      <8> <9> <10> <12> gen * bbmsl
      <10.95> bbmsl10%
      <14.4> <17.28> <20.74> <24.88> bbmsl12}{}
\DeclareFontShape{U}{bbm}{bx}{n}
   {  <5> <6> <7> <8> <9> <10> <12> gen * bbmbx
      <10.95> bbmbx10%
      <14.4> <17.28> <20.74> <24.88> bbmbx12}{}
\DeclareFontShape{U}{bbm}{bx}{sl}
   {  <5> <6> <7> <8> <9> <10> <10.95> <12> <14.4> <17.28>%
      <20.74> <24.88> bbmbxsl10}{}
\DeclareFontShape{U}{bbm}{b}{n}
   {  <5> <6> <7> <8> <9> <10> <10.95> <12> <14.4> <17.28>%
      <20.74> <24.88> bbmb10}{}
\DeclareMathAlphabet{\mathbbm}{U}{bbm}{m}{n}
\SetMathAlphabet\mathbbm{bold}{U}{bbm}{bx}{n}

\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}
\newcommand{\xmark}{\ding{55}}
\DeclareSymbolFont{extraup}{U}{zavm}{m}{n}
\DeclareMathSymbol{\varheart}{\mathalpha}{extraup}{86}

\graphicspath{{images/}}

\usepackage[protrusion=true,expansion=true]{microtype}

\setlength\parskip{\medskipamount}
\setlength\parindent{0pt}

\title{Modal operators for a constructive account of well quasi-orders}

\author{Ingo Blechschmidt}
\date{November 30th, 2024}

%\setbeameroption{show notes on second screen=bottom}
\newcommand{\jnote}[2]{\only<#1>{\note{\setlength\parskip{\medskipamount}\footnotesize\justifying#2\par}}}

%\useinnertheme[shadow=true]
\setbeamerfont{block title}{size={}}

\useinnertheme{rectangles}

\usecolortheme{orchid}
\usecolortheme{seahorse}
\definecolor{mypurple}{RGB}{253,73,34}
\definecolor{mypurpledark}{RGB}{100,0,150}
\setbeamercolor{structure}{fg=mypurple}
\setbeamercolor*{title}{bg=mypurple,fg=white}
\setbeamercolor*{titlelike}{bg=mypurple,fg=white}
\setbeamercolor{frame}{bg=black}

\usefonttheme{serif}
\usepackage[T1]{fontenc}
\usepackage{libertine}

% lifted from https://arxiv.org/abs/1506.08870
\DeclareFontFamily{U}{min}{}
\DeclareFontShape{U}{min}{m}{n}{<-> udmj30}{}
\newcommand\yon{\!\text{\usefont{U}{min}{m}{n}\symbol{'210}}\!}

\newcommand{\A}{\mathcal{A}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\C}{\mathcal{C}}
\newcommand{\M}{\mathcal{M}}
\renewcommand{\AA}{\mathbb{A}}
\newcommand{\BB}{\mathbb{B}}
\newcommand{\pp}{\mathbbm{p}}
\newcommand{\MM}{\mathbb{M}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\FF}{\mathbb{F}}
\newcommand{\G}{\mathcal{G}}
\newcommand{\J}{\mathcal{J}}
\newcommand{\GG}{\mathbb{G}}
\renewcommand{\O}{\mathcal{O}}
\newcommand{\K}{\mathcal{K}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\TT}{\mathbb{T}}
\newcommand{\PP}{\mathbb{P}}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\CC}{\mathbb{C}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\aaa}{\mathfrak{a}}
\newcommand{\bbb}{\mathfrak{b}}
\newcommand{\ccc}{\mathfrak{c}}
\newcommand{\ppp}{\mathfrak{p}}
\newcommand{\fff}{\mathfrak{f}}
\newcommand{\mmm}{\mathfrak{m}}
\newcommand{\defeq}{\vcentcolon=}
\newcommand{\defeqv}{\vcentcolon\equiv}
\newcommand{\Cov}{\mathrm{Cov}}
\renewcommand{\_}{\mathpunct{.}}
\newcommand{\?}{\,{:}\,}
\newcommand{\speak}[1]{\ulcorner\text{\textnormal{#1}}\urcorner}
\newcommand{\inv}{inv.\@}
\newcommand{\forces}{\vDash}
\newcommand{\ind}{\ensuremath{_\text{ind}}\xspace}
\newcommand{\sinf}{\ensuremath{_\infty}\xspace}
\newcommand{\impl}{\ensuremath{_\text{impl}}\xspace}

\setbeamertemplate{blocks}[shadow=false]

\newenvironment{indentblock}{%
  \list{}{\leftmargin\leftmargin}%
  \item\relax
}{%
  \endlist
}

% Adapted from https://latex.org/forum/viewtopic.php?t=2251 (Stefan Kottwitz)
\newenvironment<>{hilblock}{
  \begin{center}
    \begin{minipage}{9.05cm}
      \setlength{\textwidth}{9.05cm}
      \begin{actionenv}#1
        \def\insertblocktitle{}
        \par
        \usebeamertemplate{block begin}}{
        \par
        \usebeamertemplate{block end}
      \end{actionenv}
    \end{minipage}
  \end{center}}

\newenvironment{changemargin}[2]{%
  \begin{list}{}{%
    \setlength{\topsep}{0pt}%
    \setlength{\leftmargin}{#1}%
    \setlength{\rightmargin}{#2}%
    \setlength{\listparindent}{\parindent}%
    \setlength{\itemindent}{\parindent}%
    \setlength{\parsep}{\parskip}%
  }%
  \item[]}{\end{list}}

\tikzset{
  invisible/.style={opacity=0,text opacity=0},
  visible on/.style={alt={#1{}{invisible}}},
  alt/.code args={<#1>#2#3}{%
    \alt<#1>{\pgfkeysalso{#2}}{\pgfkeysalso{#3}}}
}

% https://tex.stackexchange.com/questions/172336/drawing-roman-laurel-leaves-spqr-in-tikz
\tikzset{
  laurel-wreath/.pic = {
    \fill svg{M14.4-24.6c-1.5-1.5-2.6-3.3-3.1-5.3l-.4-1.7c-.2-1.1-.2-4.1 .2-5.7 .2-.9 .3-1.3 .5-1.3l1.4 1.1 2.5 2.4c2.7 2.5 5.2 6 5.8 8 .2 .6-.5 .3-2.2-.9-1.6-1.3-3.3-2.6-5-3.8l.1 1.4c.2 1.4 .5 2.7 1.1 4.6s.8 2.5 .5 2.5l-1.4-1.3zm69.6 1.1 .3-1.2c.8-2.3 1.3-4.8 1.6-7.3l-1.5 1.1c-1.3 .9-2.6 1.9-3.7 3-1.6 1.1-2 1.3-2.1 1 .7-1.8 1.6-3.4 2.8-4.9 1.3-1.7 6.5-6.8 7-6.8 .2 0 .3 .2 .3 .5l.3 1.6c.3 2.2 .2 5.7-.5 7.4-.8 1.9-1.6 3.1-3 4.7-1.1 1.1-1.4 1.3-1.5 .9z};
    \fill svg{M10-29.4c-.8-1.1-1.4-2.2-2-4.1l-.7-3.5c-.2-3 .2-4.4 1.4-8.3l.5-1.4c.2-1.3 .3-1.9 .6-1.9 .3-.2 .6 .3 .7 .8s.9 2.2 1.9 3.6c1.4 2.2 2.7 4.4 3.9 6.6l.9 2.7c0 .6 0 .6-.3 .6-.6 0-4.9-4.4-5.8-6l-.2-.6-.1 1.7-.3 2.8c-.3 2.7-.3 3.8 0 5.5 .6 2 .5 2.4-.5 1.5zm79.2 .3 .4-2.4c.2-1.3 .2-2.7-.1-4.9l-.3-2.8v-1.6l-.7 1c-.8 1.3-5 5.5-5.5 5.5s-.5-.3 .2-1.9c.5-1.7 1.4-3.3 3.3-6.5 2.4-3.6 2.7-3.9 2.8-4.7 .5-1.3 .5-1.4 .8-1.2 .3 0 .6 .8 .6 1.5l.7 2.4c.9 2.7 1.1 3.6 1.2 6 .2 3.1-.5 6-2 8.2-.8 1.3-1.3 1.7-1.4 1.5z};
    \fill svg{M5-40c-.4-3.2-.1-6.5 .9-9.6 .5-1.1 1.6-2.8 2.2-3.4l1.3-1.6 2-2.7 .2 .6c.1 1.3 .4 2.6 .9 3.8l.3 1c.8 1.7 1.1 2.7 1.6 5.3 .6 2.5 .6 4.6 .2 4.6-.3 0-.9-.8-1-1.1l-.5-.8c-1.4-2-3-5.2-2.9-6.5-.9 2.7-2 5.4-3.5 7.9l-.3 .8-.3 .8c0 .5-.6 1.6-.8 1.6l-.3-.7zm89.2 .2-.2-.5-.3-.9-1.1-2.7-1.1-2.4c-.6-1.4-1.2-2.8-1.6-4.2l-.3 .9c-.3 1.3-1.6 3.9-3 6-1.3 2-1.6 2-1.5 0s1.1-6.3 2.2-9c.8-1.7 1.1-3.1 .9-4.1-.2-1.1 .5-.8 2.2 1.8 3.3 4.4 3.8 5.4 4.4 7.8 .6 2.4 .5 7.7-.3 7.8l-.3-.5z};
    \fill svg{M13.9-50.1c-.5-1.9-.8-3.9-.9-5.8-.2-1.6-.1-3.3 .1-4.9-.3 .8-1.7 2.5-4.2 5.1l-3 4.9-.3 .1c-.3 0-.3-2.2 0-3.3 .8-3 1.4-4.6 2.5-6.1 .9-1.3 1.7-1.9 2.5-2.5 1.1-.6 2.7-1.9 3.5-2.7 .9-.9 1.9-1.4 2.2-1.4v1.1l-.3 6.6c0 6.8 .2 6.3-1 8.9-.5 1.1-.8 1.1-1.1 0zm70.8-.4c-.8-2.2-.8-2.5-.7-6.3-.1-2.7-.1-5.5-.2-8.2-.3-1.6-.3-1.9 .5-1.6l.6 .5c1.4 1.4 3 2.5 3.9 3.1 1.3 .9 1.9 1.6 2.7 2.6l.6 .7 .2 .4 .2 .3c.8 .9 2 4.9 2 6.9 .2 1.9-.2 1.9-.9 .5-.7-1.4-1.5-2.7-2.6-4-1.6-1.5-3-3.2-4.2-5 .4 3 .3 6-.5 9 0 .8-.5 2.2-.8 2.3-.2 0-.5-.3-.8-1.2z};
    \fill svg{M16.4-58.5l.2-1.5 .3-3.7c.2-2.8 .3-3.5 1.1-5.4l.7-1.3-.5 .4-1 .7c-.5 .4-1.1 .8-1.5 1.3l-.5 .3-1.9 1.6c-2.2 1.6-2.7 2-3.9 3.6-.5 .8-1.1 1.3-1.3 1.3-.5 0 0-2.4 1.1-4.7 1.5-3.4 4.3-6 7.7-7.4l1.3-.4 1.9-.4 2-.5c1.4 0 1.4 0 1 1.1-.5 .8-.8 2-1.1 4.2-.3 2.3-1.1 4.5-2.2 6.5l-.4 .6c-.6 1.1-1.3 2.1-2 3.2-.5 .6-.8 .8-1 .5zm66.3-.2c-.8-.9-2.8-4.4-3.5-6.1-.6-1.3-.9-2.5-1.1-3.5-.2-2.1-.7-4.1-1.5-6 0-.3 0-.3 1.2-.3l2.1 .5 1.9 .4 1.2 .4 .6 .1 1 .6c3 1.4 5.7 4.6 6.8 8.5l.7 2.6c-.2 .6-.5 .5-1.4-.7-2.2-2.7-4.8-5-7.7-6.9l-1.7-1.3 .6 1.3c.3 .6 .6 1.2 .8 1.9l.3 2.5 .3 3.9c.3 2.4 .2 2.8-.6};
    \fill svg{M21.6-66.1l.4-1.1 .9-3.2c.3-1.9 1.1-3.3 2.4-4.7l.4-.8-1.2 .2-2.2 .3c-2.7 .3-5.3 1.2-7.7 2.5-.6 .5-1.3 .6-1.3 .3 0-.5 .9-1.9 2-2.9 .8-.9 2-1.9 3.2-2.6l.9-.4 2.2-1c.3-.2 1.3-.3 3.2-.1 3 0 4.1 .2 6.3 .7l1.1 .4c.5 .2 .6 .6 .3 .6-.5 0-1.4 .9-1.9 1.7l-1.2 1.8c-1.7 2.8-2.2 3.5-4.6 5.9l-3 2.7-.2-.3zm53.9-2c-2.7-2.8-3.5-3.8-5.4-6.8-.9-1.6-1.4-2.4-1.9-2.5l-.8-.5c-.3 0-.2-.5 .4-.6l1.1-.4c1.9-.6 3-.8 5.6-.9l3.3 .2c2 .6 3.8 1.5 5.4 2.8 .3 0 1.9 1.6 2.5 2.4l.9 1.8c0 .3-.3 .2-1.9-.6-2.8-1.4-4.4-1.9-7.7-2.2l-2.2-.5c-.9-.2-.9-.2-.6 .2 .6 .5 1.7 2 2.1 2.8l.9 2.5c.3 1.5 .6 3 .9 4.6l-2.6-2.3z};
    \fill svg{M34.1-78.7c-3.4-1.3-6.9-2.1-10.6-2.5-.9 0-1.4 0-2.3 .3-2 .5-2 0 0-1.3l2.8-1.2c1.4-.5 1.9-.5 3.8-.6 3.8-.2 6.1 .3 9.3 1.7l3.6 1.1 2.2 .3c1.3 0 1.7 0 2.7-.3 1.1-.3 2.8-1.1 2.8-1.3l-1.3-.9c-1.9-1.4-3.1-2.7-3.1-3.2l.8-.6c.9-.3 1.3-.2 2 .8 .5 .8 1.1 1.4 2.9 2.7 .2 .3 .3 .2 1.1-.3 .9-.8 2.4-2 2.6-2.7 .5-.6 .9-.8 1.8-.5l.8 .6c0 .5-1.4 1.7-3.2 3.2l-1.3 .9c0 .2 1.7 .9 2.9 1.3 .9 .3 1.4 .3 2.7 .3l2.2-.3c1.7-.4 3.4-1 5-1.7 2-.8 4.4-1.3 7.7-1.1 2 .2 2.5 .2 3.8 .6 .9 .3 2.2 .8 2.8 1.2 2 1.1 2 1.6 .2 1.3-1.6-.3-1.9-.3-4.4 0-2.4 .3-4.7 .8-7 1.6l-1.5 .6c-2.9 .3-5.9 .2-8.8-.3-1.7-.3-3.6-.9-6-2.1l-1.1-.4-1.3 .6c-4.5 2.2-9.6 3-14.6 2.2zm-6.3-9.1c};
  }
}

\newcommand{\pointthis}[3]{%
  \tikz[remember picture,baseline]{
    \node[anchor=base,inner sep=0,outer sep=0] (#2) {#2};
    \node[visible on=#1,overlay,rectangle callout,rounded corners,callout relative pointer={(0.3cm,0.5cm)},fill=blue!20] at ($(#2.north)+(-0.1cm,-1.1cm)$) {#3};
  }%
}

\tikzset{
  invisible/.style={opacity=0,text opacity=0},
  visible on/.style={alt={#1{}{invisible}}},
  alt/.code args={<#1>#2#3}{%
    \alt<#1>{\pgfkeysalso{#2}}{\pgfkeysalso{#3}}}
}

\newcommand{\hcancel}[5]{%
  \tikz[baseline=(tocancel.base)]{
    \node[inner sep=0pt,outer sep=0pt] (tocancel) {#1};
    \draw[red!80, line width=0.4mm] ($(tocancel.south west)+(#2,#3)$) -- ($(tocancel.north east)+(#4,#5)$);
  }%
}

\newcommand{\explain}[7]{%
  \tikz[remember picture,baseline]{
    \node[anchor=base,inner sep=2pt,outer sep=0,fill=#3,rounded corners] (label) {#1};
    \node[anchor=north,visible on=<#2>,overlay,rectangle callout,rounded corners,callout
    relative pointer={(0.0cm,0.5cm)+(0.0cm,#6)},fill=#3] at ($(label.south)+(0,-0.3cm)+(#4,#5)$) {#7};
  }%
}

\newcommand{\explainstub}[2]{%
  \tikz[remember picture,baseline]{
    \node[anchor=base,inner sep=2pt,outer sep=0,fill=#2,rounded corners] (label) {#1};
  }%
}

\newcommand{\squiggly}[1]{%
  \tikz[remember picture,baseline]{
    \node[anchor=base,inner sep=0,outer sep=0] (label) {#1};
    \draw[thick,color=red!80,decoration={snake,amplitude=0.5pt,segment
    length=3pt},decorate] ($(label.south west) + (0,-2pt)$) -- ($(label.south east) + (0,-2pt)$);
  }%
}

% Adapted from https://latex.org/forum/viewtopic.php?t=2251 (Stefan Kottwitz)
\newenvironment<>{varblock}[2]{\begin{varblockextra}{#1}{#2}{}}{\end{varblockextra}}
\newenvironment<>{varblockextra}[3]{
  \begin{center}
    \begin{minipage}{#1}
      \begin{actionenv}#4
        {\centering \hil{#2}\par}
	\def\insertblocktitle{}%\centering #2}
        \def\varblockextraend{#3}
	\usebeamertemplate{block begin}}{
        \par
        \usebeamertemplate{block end}
        \varblockextraend
      \end{actionenv}
    \end{minipage}
  \end{center}}

\setbeamertemplate{headline}{}

\setbeamertemplate{frametitle}{%
  \leavevmode%
  \vskip-1.6em%
  \begin{beamercolorbox}[dp=1ex,center,wd=\paperwidth,ht=2.25ex]{title}%
    \vskip0.5em%
    \bf\insertframetitle
  \end{beamercolorbox}%

  \vskip-0.77em\hspace*{-2em}%
  \textcolor{mypurpledark}{\rule[0em]{1.1\paperwidth}{2.4pt}}

  \vskip-0.4em%
}

\setbeamertemplate{navigation symbols}{}

\newcounter{framenumberpreappendix}
\newcommand{\backupstart}{
  \setcounter{framenumberpreappendix}{\value{framenumber}}
}
\newcommand{\backupend}{
  \addtocounter{framenumberpreappendix}{-\value{framenumber}}
  \addtocounter{framenumber}{\value{framenumberpreappendix}}
}

\newcommand{\insertframeextra}{}
\setbeamertemplate{footline}{%
  \begin{beamercolorbox}[wd=\paperwidth,ht=2.25ex,dp=1ex,right,rightskip=1mm,leftskip=1mm]{}%
    % \inserttitle
    \hfill
    \insertframenumber\insertframeextra\,/\,\inserttotalframenumber
  \end{beamercolorbox}%
  \vskip0pt%
}

\newcommand{\hil}[1]{{\usebeamercolor[fg]{item}{\textbf{#1}}}}
\newcommand{\hill}[1]{{\usebeamercolor[fg]{item}{#1}}}
\newcommand{\bad}[1]{\textcolor{red!90}{\textnormal{#1}}}
\newcommand{\good}[1]{\textcolor{mypurple}{\textnormal{#1}}}

\newcommand{\bignumber}[1]{%
  \renewcommand{\insertenumlabel}{#1}\scalebox{1.2}{\!\usebeamertemplate{enumerate item}\!}
}
\newcommand{\normalnumber}[1]{%
  {\renewcommand{\insertenumlabel}{#1}\!\usebeamertemplate{enumerate item}\!}
}
\newcommand{\bigheart}{\includegraphics{heart}}

\newcommand{\subhead}[1]{{\centering\textcolor{gray}{\hrulefill}\quad\textnormal{#1}\quad\textcolor{gray}{\hrulefill}\par}}

\newcommand{\badbox}[1]{\colorbox{red!30}{#1}}
\newcommand{\infobox}[1]{\colorbox{yellow!70}{\color{black}#1}}

% taken from JDH "The modal logic of arithmetic potentialism and the universal algorithm"
\DeclareMathOperator{\possible}{\text{\tikz[scale=.6ex/1cm,baseline=-.6ex,rotate=45,line width=.1ex]{\draw (-1,-1) rectangle (1,1);}}}
\DeclareMathOperator{\necessary}{\text{\tikz[scale=.6ex/1cm,baseline=-.6ex,line width=.1ex]{\draw (-1,-1) rectangle (1,1);}}}
\DeclareMathOperator{\xpossible}{\text{\tikz[scale=.6ex/1cm,baseline=-.6ex,rotate=45,line width=.1ex]{\draw (-1,-1) rectangle (1,1); \draw[very thin] (-.6,-.6) rectangle (.6,.6);}}}
\DeclareMathOperator{\xnecessary}{\text{\tikz[scale=.6ex/1cm,baseline=-.6ex,line width=.1ex]{\draw (-1,-1) rectangle (1,1); \draw[very thin] (-.6,-.6) rectangle (.6,.6);}}}

\input{images/primes.tex}

\newcommand{\triang}{\hil{$\blacktriangleright$}}
\newcommand{\concat}{\mathbin{{+}\mspace{-8mu}{+}}}

\newcommand{\astikznode}[2]{\tikz[baseline,remember picture]{\node[anchor=base,inner sep=0,outer sep=0.1em] (#1) {#2};}}
\newcommand{\astikznodecircled}[3]{\tikz[baseline,remember picture]{\node[anchor=base,circle,draw=#2,thick,inner sep=0.05em,outer sep=0.05em] (#1) {#3};}}
\newcommand{\astikznodetransparentlycircled}[2]{\tikz[baseline,remember picture]{\node[anchor=base,circle,opacity=0,draw=white,text opacity=1,thick,inner sep=0.05em,outer sep=0.05em] (#1) {#2};}}

\setbeamersize{text margin left=1.60em,text margin right=1.60em}

\newlength\stextwidth
\newcommand\makesamewidth[3][c]{%
  \settowidth{\stextwidth}{#2}%
  \makebox[\stextwidth][#1]{#3}%
}

\newcommand{\dnote}[1]{%
  \begin{tabular}{@{}m{2em}@{}m{0.83\textwidth}@{}}%
    \textdbend &#1%
  \end{tabular}%
  \par
}

\newcommand{\genalpha}{\mbox{$\hspace{0.12em}\shortmid\hspace{-0.62em}\alpha$}}

\usepackage{newunicodechar}

\newunicodechar{∇}{\ensuremath{\nabla}}
\newunicodechar{λ}{\ensuremath{\lambda}}
\newunicodechar{σ}{\ensuremath{\sigma}}
\newunicodechar{τ}{\ensuremath{\tau}}
\newunicodechar{∷}{\ensuremath{\!\!\!}}
\newunicodechar{⧺}{\ensuremath{\!\!\!}}
\newunicodechar{₁}{\ensuremath{_1}}
\newunicodechar{₂}{\ensuremath{_2}}
\newunicodechar{∈}{\ensuremath{\in}}

\begin{document}

\addtocounter{framenumber}{-1}

\definecolor{mypurpleblack}{RGB}{30,0,50}
\setbeamercolor{structure}{fg=black}

{\usebackgroundtemplate{\begin{minipage}{\paperwidth}\centering\vspace*{-7em}\includegraphics[width=\paperwidth]{forest2}\end{minipage}}
\begin{frame}[c]
  \centering
  \color{white}

  \bigskip
  \bigskip
  \bigskip
  \bigskip
  \bigskip
  \bigskip
  \bigskip
  \bigskip
  \bigskip

  \scriptsize

  \setbeamercolor{block body}{bg=black!100}
  \begin{minipage}{0.52\textwidth}
    \begin{block}{}
      \centering\normalsize\color{white}
      \hil{\color{white}Towards topological type theory for decrypting
      transfinite methods in classical~mathematics} \\[-0.4em]
      \
    \end{block}
  \end{minipage}

  \bigskip
  \bigskip
  \bigskip
  \bigskip
  \bf
  \colorbox{black}{\begin{minipage}{0.2\textwidth}
    \centering
    TYPES 2025 \\
    June 9th, 2025
  \end{minipage}}
  \bigskip

  \colorbox{black}{\begin{minipage}{0.2\textwidth}
    \centering
    Ingo Blechschmidt \\
    University of Antwerp
  \end{minipage}}
\end{frame}}

\definecolor{mypurple}{RGB}{150,0,255}
\setbeamercolor{structure}{fg=mypurple}

{\usebackgroundtemplate{\begin{minipage}{\paperwidth}\vspace*{1.7em}\includegraphics[height=\paperheight]{sea-of-clouds-1}\end{minipage}}
\begin{frame}{A case study in Hilbert's program}
  \textbf{Def.} Let~$(X,{\leq})$ be a quasi-order.
  \vspace*{-0.3em}
  \begin{itemize}
    \item A sequence~$\alpha : \NN \to X$ is \hil{good} iff
    there merely exist~$i < j$ with~$\alpha\,i \leq \alpha\,j$.
    \item The quasi-order~$X$ is \hil{well\only<4->{\sinf}} iff every sequence~$\NN \to X$ is good.
  \end{itemize}
  \pause
  \vspace*{-1.3em}

  \begin{columns}[t]
    \begin{column}{0.46\textwidth}
      \begin{block}{Natural numbers\phantom{y}}
        \textbf{Prop.} $(\NN, {\leq})$ is well\only<4->{\sinf}.
        \smallskip

        \emph{Proof.} Let~$\alpha : \NN \to \NN$. By~\badbox{\textsc{lem}}, there is a
        \bad{minimum}~$\alpha\,i$.
        Set~$j \defeq i + 1$. \qed

        \hfill\color{gray}offensive?
      \end{block}
      \only<2-6>{
        \vspace*{-2em}
        \[ \astikznodetransparentlycircled{xm}{7}\!,
        \quad \astikznodetransparentlycircled{x0}{4}\!,
        \quad \astikznodecircled{t1}{mypurple}{3}\!,
        \quad \astikznodetransparentlycircled{x1}{1}\!,
        \quad \astikznodecircled{t2}{mypurple}{8}\!,
        \quad \astikznodetransparentlycircled{x2}{2}\!,
        \quad \ldots \]
      {\centering\begin{tikzpicture}[remember picture,overlay]
        \node[draw=mypurple, circle, thick, inner sep=0.1em] (t3) {\scriptsize$\leq$};
        \path[draw=mypurple,thick]
          (t1)
          to [out=-90, in=180] (t3)
          to [out=0, in=-90] (t2);
      \end{tikzpicture}\par}}
    \end{column}

    \pause

    \begin{column}{0.50\textwidth}
      \begin{block}{Key stability results}
        \justifying
        \only<1-5>{Assuming~\badbox{\textsc{lem}}
        and~\badbox{\textsc{dc}}}\only<6->{\good{Constructively}}, \ldots

        \hil{Dickson:} \tabto{1.8cm} If~$X$ and~$Y$ are well\only<4-5>{\sinf}\only<6->{\ind}, so is~$X \times Y$. \\
        \hil{Higman:}  \tabto{1.8cm} If~$X$ is well\only<4-5>{\sinf}\only<6->{\ind}, so is~$\mathrm{List}\,X$. \\
        \hil{Kruskal:} \tabto{1.8cm} If~$X$ is well\only<4-5>{\sinf}\only<6->{\ind}, so is~$\mathrm{Tree}\,X$.
      \end{block}
    \end{column}
  \end{columns}
  \pause
  \pause
  \pause
  \pause

  \textbf{Def.} A quasi-order~$X$ is \hil{well\ind} iff~$G\,[\,]$, where $G$ is
  the following inductively defined predicate on \hil{finite lists}.
  (In presence of \bad{bar induction}, $\text{well}\ind \Leftrightarrow
  \text{well}\sinf$.)
  \[
    \small
    \infer{\mathsf{now}\,p : G\,\sigma}{p : \mathsf{Good}\,\sigma}
    \qquad
    \infer{\mathsf{later}\,f : G\,\sigma}{f : (x : X) \to G\,(\sigma ::^r x)}
  \]
  \vspace*{-1.3em}

  \only<8>{
    \begin{columns}[c]
      \begin{column}{0.01\textwidth}
	\includegraphics[height=2.4em]{question-mark}
      \end{column}
      \begin{column}{0.9\textwidth}
      Is there a procedure for reinterpreting \hil{classical proofs}
      regarding well\sinf{} as \newline \hil{blueprints for constructive proofs} regarding well\ind?
      \end{column}
    \end{columns}
  }

  \only<9>{
    \vspace*{-0.0em}
    \hil{Central insight:} A quasi-order $X$ is well\ind iff $\hil{$\necessary$}\,\forall\alpha : \NN \to
    X\_ \exists i < j\_ \alpha\,i \leq \alpha\,j$.
  }
\end{frame}}

\begin{frame}{Missing functions in the type of all functions?}
  \begin{columns}[t]
    \begin{column}{0.57\textwidth}
      \begin{block}{Behold: A transfinite tool \ldots\phantom{g}}
        \justifying
        \textbf{Lemma.} \badbox{\textsc{lem}} Let~$X$ be well\sinf.
        Let~$\alpha : \NN \to X$. Then there is an increasing
        subsequence~$\alpha\,i_0 \leq \alpha\,i_1 \leq \cdots$.
      \end{block}

      \small
      \emph{Proof.}
      \begin{changemargin}{-0.8em}{0em}\begin{enumerate}\justifying
      \item The type
      $I \defeq \sum_{i \? \NN} \neg\sum_{j \? \NN} i < j \times \alpha\,i \leq
      \alpha\,j$
      is not in bijection with~$\NN$, as else the~$I$-extracted subsequence
      of~$\alpha$ would not be good.\smallskip
      \item By~\badbox{\textsc{lem}}, the type~$I$ is finite.\smallskip
      \item Any index~$i_0$ larger than all the indices in~$I$ is a suitable
      starting point for an increasing subsequence. \qed
      %  \item Let $\mathsf{Sad}\,i \defeq
      %  \neg \mathsf{Happy}\,i$ be the proposition
      %  expressing that~$i$ does not appear as the first component
      %  of a good pair, where
      %  $\mathsf{Happy}\,i \defeq \sum_{j \? \NN} i < j \times \alpha\,i \leq
      %  \alpha\,j$.
      %  \item If~$\mathsf{Sad}$ is anonymously dense in~$\NN$, i.e. $\forall a
      %  \? \NN\_ \neg\neg\exists b\?\NN\_ b \geq a \times \mathsf{Sad}\,b$, then
      %  \badbox{\textsc{lem}} and \badbox{\textsc{dc}} provide us with a
      %  subsequence~$\alpha\,k_0 \leq \alpha\,k_1 \leq \ldots$
      %  with~$\mathsf{Sad}\,k_0, \mathsf{Sad}\,k_1, \ldots$. This is a contradiction to wellness.
      %  \item
      %  Hence there is~\bad{not~not} an index~$a \? \NN$ such that~$\mathsf{Sad}$ is
      %  strictly bounded by~$a$; hence every index~$b \geq a$ does~\bad{not not} have
      %  a partner: $\forall b\geq a\_ \neg\neg\exists c>b\_
      %  \alpha\,b \leq \alpha\,c$. Thus~\bad{\textsc{lem}} and \bad{\textsc{dc}}
      %  provide us with the desired subsequence. \qed
      \end{enumerate}\end{changemargin}
    \end{column}

    \begin{column}{0.43\textwidth}
      \begin{block}{\ldots{} implying a concrete consequence}
        \textbf{Cor.} \badbox{\textsc{lem}}
        Let~$X$ and~$Y$ be well\sinf. Then~$X \times Y$ is
        well\sinf.\phantom{g}
      \end{block}

      \small
      \emph{Proof.}
      \begin{changemargin}{-0.8em}{0em}\begin{enumerate}\justifying
      \item Let a sequence~$\gamma : \NN \to X \times Y$ be
      given. Write~$\gamma\,k = (\alpha\,k,\beta\,k)$.\smallskip
      \item By the lemma, there is an
      increasing subsequence~$\alpha\,i_0 \leq \alpha\,i_1 \leq \cdots$.\smallskip
      \item Because~$Y$
      is well, there are indices~$n < m$ such that~$\beta\,i_n \leq
      \beta\,i_m$.\smallskip
      \item As
      also~$\alpha\,i_n \leq \alpha\,i_m$, the sequence~$\gamma$ is good. \qed
      \end{enumerate}\end{changemargin}
    \end{column}
  \end{columns}

  \dnote{\emph{We cannot trust~\bad{\textsc{lem}}-provided sequences to be
  available in the type~$\NN \to X$. Similarly with~\bad{\textsc{dc}}.}}
\end{frame}

{\usebackgroundtemplate{\begin{minipage}{\paperwidth}\vspace*{0.29cm}\hspace*{-1cm}\includegraphics[height=\paperheight]{tower-cropped}\end{minipage}}
\begin{frame}{Where do many cherished inductive definitions come from?}
  \begin{changemargin}{8.0em}{-0.5em}
    In mathematics, we routinely enlarge structures:
    \begin{itemize}
      \small
      \item Pass from~$\mathbb{Q}$ to~$\mathbb{R}$, to embrace irrationals.
      \item Pass from~$\mathbb{R}$ to~$\mathbb{C}$, to obtain $\sqrt{-1}$.
      \item Pass from~$\mathbb{C}$ to~$\mathbb{C}[X]$, to obtain a ``generic number''.
    \end{itemize}

    In set and type theory, we can also enlarge \hil{the mathematical universe}:
    \begin{itemize}
      \small
      \item Force a \hil{generic sequence} $\NN \to X$.
      \item Force a \hil{generic enumeration} $\NN \twoheadrightarrow X$
            (even if $X$ is uncountable).
      \item Force a \hil{generic prime ideal} of a given ring.
    \end{itemize}
    \pause

    Central observations of the multiversal yoga:
    \begin{itemize}
      \small
      \item A quasi-order~$X$ is well\ind iff the generic sequence~$\NN \to X$
      is good. % iff \hil{everywhere,} every sequence~$\NN \to X$ is good.
      \item A relation is well-founded iff for the generic descending chain, $\bot$.
      \item A ring element is nilpotent iff it is contained in the generic
      prime ideal.
    \end{itemize}
    \pause
